\documentclass[10pt,letterpaper]{article}
\usepackage[right=20mm,left=20mm,top=20mm,bottom=20mm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage[utf8x]{inputenc}
\usepackage{graphicx}
\graphicspath{{../../results/}}
\usepackage{subcaption}
\usepackage{caption}
\usepackage{braket}

\begin{document}
	
    \LARGE{\textbf{Research log summary}}\hfill\Large{June 8, 2020}
    \section*{Progress}
		\begin{itemize}
			\item Imposing tradeoff makes dynamics more complex. Probably because now reaction networks are more diverse (both specialists and generalists are equally fit).
			\item  $ \braket{f} $ is no longer a good predictor of the performance at community even for low values of richness.
			\item Created community reaction network, with m nodes and $ n_{reac} $ edges. Calculated betweenness, in-degree, out-degree, and their respective standar deviations. These would measure how efficiently the community depleates resources simultaneously.
			\item Betweenness doesn't seem to be a quantity of interest, since it doesn't display recognizable patterns. 
			\item Read a lot. Defined my thesis more precisely
    	\end{itemize}
    \section*{Questions}
	\begin{itemize}
	   \item I want to investigate 4 main questons:\\
	         1. Have repeated and continuous coalescence events contributed to the high microbial diversity in some habitats, such as the soil?\\
	        2. Will members of communities with a history of coalescence have a higher persistence up on interaction with a 'naive' community?\\
	        3. Can we predict properties of coalescent communities using approaches derived from the study of interacting networks?\\
	        In which regime should I study this questions: diversification regime, or all resources present regime?
		\item I am setting the numer of reactions of each strain by choosing a number from a uniform distribution [1, ${m}\choose{2}$]. Maybe this is not realistic, because it gnereates network with a very high connectance (more than realistically feasible).
		\item Record modularity?
	
	\end{itemize}

	\section*{Next steps}
		\begin{itemize}
			\item Start performning community coalescence events and record all the quantities (richness, degree of network, richness, abundances) in $ C_1 $, $ C_2 $, $ C $. 
			\item Keep reading communiy coalescence papers.
			\item Vary environment conditions (diversification/ all resources)\\
		\end{itemize}
	
		
		
\end{document}
